Hunting the Parent of the Orphan Stream: Identifying Stream Members \\ from Low-Resolution Spectroscopy

### 2013-02-27 thesis/chapter2/chapter2.tex
Added thesis as of 27 Feb

\documentclass[10pt,apjl]{emulateapj}
\usepackage{amssymb,graphics} % for resolution symbol
\usepackage{multirow} % deluxetable wasn't doing it for me.

\begin{document}

\shorttitle{Identifying Orphan Stream Members from Low-Resolution Spectroscopy}
\shortauthors{Casey et al}

\title{Hunting the Parent of the Orphan Stream: Identifying Stream Members \\ from Low-Resolution Spectroscopy}

\author{Andrew R. Casey\altaffilmark{1,2}, Gary Da Costa\altaffilmark{1}, Stefan C. Keller\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Massachusetts Institute of Technology and Kavli Institute for
Astrophysics and Space Research, 77 Massachusetts Avenue,
Cambridge, MA 02139, USA}

\begin{abstract}
We present candidate K-giant members in the Orphan Stream which have been identified from low-resolution data taken with the AAOmega spectrograph on the Anglo-Australian Telescope. From modest S/N spectra and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. We find a revised stream distance of $22.5\,\pm\,2.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 82.1\,\pm\,1.4$\,km s$^{-1}$. The observed velocity dispersion of our most probable members is consistent with arising from the velocity uncertainties alone. This indicates that at least along this line-of-sight, the Orphan Stream is kinematically cold. Our data indicates an overall stream metallicity of [Fe/H] $= -1.63\,\pm\,0.19$\,dex which is more metal-rich than previously found and unbiased by spectral type. Furthermore, the significant metallicity dispersion displayed by our most probable members, $\sigma(\mbox{[Fe/H]}) = 0.56$\,dex, suggests that the unidentified Orphan Stream parent is a dSph satellite. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}

@@

In this paper we present a detailed, self-consistent analysis to identify K-giant members of the Orphan Stream. Using our selection method we have catalogued the locations of nine highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:analysis} and in \S\ref{sec:conclusions} the conclusions, predictions and future work are presented.

\section{Target Selection}
\label{sec:target-selection}

@@

\section{Observations}
\label{sec:observations}

@@

\section{Analysis \& Discussion}
\label{sec:analysis}

@@

Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin_et-al_2003}. We have selected particles from the Besan\c{c}on model using the same criteria outlined in \S\ref{sec:target-selection} after employing the \citet{Jordi_et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$.

@@

In a nearby ($\Delta\Lambda_{Orphan} \sim 4\,^\circ$) region of the stream, \citet{Newberg_et-al_2010} detected the Orphan Stream with a $V_{GSR} = 101.4$ km s$^{-1}$ from BHB stars. Differences in accounting for the local standard of rest between this work and \citet{Newberg_et-al_2010} means that this corresponds to approximately 95\,km s$^{-1}$ on our $V_{GSR}$ scale.  This is discussed further in \S\ref{sec:newberg}. The expected Orphan Stream kinematic peak is labelled in Figure \ref{fig:velocities}. There is no obvious sharp kinematic peak representative of the Orphan Stream in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution. To isolate potential Orphan Stream members we have nominated a relatively wide selection criteria between $65 \leq V_{GSR} \leq 125$ km s$^{-1}$ (shown in Figure \ref{fig:velocities}), which yields 28 Orphan Stream candidates. The typical uncertainty in our velocities is $\pm{}5.0$\,km s$^{-1}$.

\subsection{Dwarf/Giant Discrimination}
\label{sec:dwarf-giant}

@@

Figure \ref{fig:ew-mg} illustrates the trend with $EW_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel_Finkbeiner_Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. Giant stars populate the lower, sparser branch. A separation line has been adopted to distinguish dwarfs from giants, and is shown in Figure \ref{fig:ew-mg}. If we were to place this line higher, the total number of true giant stars may increase, but the dwarf contamination rate will rise dramatically. A compromise must be made between the rate of giant recoverability and the dwarf contamination. Our  dwarf/giant separation line lies just below the main dwarf population. On its own, this dwarf/giant separation line would typically result in far too many dwarf contaminants. However, we are  employing selections on multiple observables (kinematics, metallicity, proper motions) in order to refine our Orphan Stream giant sample.

@@

Our dwarf/giant separation line in Figure \ref{fig:ew-mg} yields 425 potential giants. Upon taking the intersection of our kinematic and gravity selections, we find 20 stars that appear to be likely Orphan Stream giants.

\subsection{Metallicities}
\label{sec:metallicities}

@@

Given a distance estimate to the Orphan Stream, we can also deduce a star's metallicity through isochrone fitting. We have used a 10\,Gyr \citet{Girardi_et-al_2008} isochrone at 21.4\,kpc \citep{Newberg_et-al_2010} and found metallicities for all 20 likely stream members from their best-fitting isochrone. Derived metallicities from Ca\,\textsc{II} line strengths and isochrone fitting that are consistent (within $\pm0.3$\,dex) indicates these measurements are reliable, and that these stars are indeed at a distance of $\sim$21.4\,kpc. We find ten highly likely stream members with consistently derived metallicities. They fall within the shaded region illustrated in Figure \ref{fig:feh}.

@@

Since our Orphan Stream giants cover a wide evolutionary range along the giant branch (Figure \ref{fig:cmd}), we are in a good position to revise the distance estimate to the stream. Given a 10\,Gyr \citet{Girardi_et-al_2008} isochrone at $[\mbox{Fe/H}] = -1.63$, we find a best-fitting distance to the stream of $22.5 \pm 2.0$\,kpc at $(l, b) = (250\,^\circ,\,50\,^\circ)$. This isochrone is shown in Figure \ref{fig:cmd}. Our derived distance is in reasonably good agreement with the measurement of $21.4 \pm 1.0$\,kpc independently deduced by \citet{Grillmair_2006} and \citet{Newberg_et-al_2010}.

\subsection{Comparison with \citet{Newberg_et-al_2010}}
\label{sec:newberg}
\citet{Newberg_et-al_2010} traced the Orphan Stream using BHB stars selected from the SEGUE survey, allowing them to derive an orbit for the stream and make a strong prediction for the location of the undiscovered progenitor. Their closest stream detection to this study is at $\Lambda_{Orphan} = 18.4\,^\circ$, approximately $\Delta\Lambda_{Orphan} \sim 4\,^\circ$ away from our fields. At this location, \citet{Newberg_et-al_2010} found the velocity of the stream to be $V_{GSR} = 101.4 \pm 8.9$\,km s$^{-1}$ based on 12 BHB stars. We note that this is $\sim95$\,km s$^{-1}$ on our scale, given the differences in accounting for the local standard of rest. The velocities and metallicities of our `High' and `Medium' probability candidates are illustrated in Figure \ref{fig:vgsr-feh}. Although we recover some candidates with velocities up to $V_{GSR} \sim 110$\,km s$^{-1}$, our kinematic distribution peaks near $V_{GSR} \sim 85$\,km s$^{-1}$, roughly 10\,km s$^{-1}$ lower than that of \citet{Newberg_et-al_2010}.

@@

The adopted metallicities of our Orphan Stream candidates are generally higher than those found by \citet{Newberg_et-al_2010}. Our highly likely stream members have a mean metallicity of [Fe/H] = $-1.63$, with a dispersion of $\sigma = 0.56$\,dex. As illustrated in Figures \ref{fig:vgsr-feh} and \ref{fig:newberg-feh}, there are two very metal-poor candidates which largely drive this dispersion, but we have no reason to suspect they are non-members. The \citet{Newberg_et-al_2010} sample contains 37 BHB stars identified over a 60\,$^\circ$ arc on the sky, and has a peak metallicity at [Fe/H] = $-2.10 \pm 0.10$. The closest detection bin in the \citet{Newberg_et-al_2010} sample was the most populous, yielding 7 BHB stars. For comparison, we identify 9 giant stars across $\sim{}4\,^\circ$. Given BHB stars are known to trace a somewhat more metal-poor population, and we are calculating statistics with marginal sample sizes, we conclude that the accuracy of these two metallicity distributions are not mutually exclusive. It is entirely possible that we are sampling the same distribution, but a larger sample size is required.

\section{Conclusions}
\label{sec:conclusions}

@@

Given the stream orbit derived by \citet{Newberg_et-al_2010}, they excluded all possible known halo objects except for the dissolved star cluster, Segue 1. \citet{Simon_et-al_2011} obtained spectroscopy for six members in Segue 1 and found an extremely wide metallicity dispersion: from $<-3.4$ to $-1.63$ dex. On the basis of the extremely low metallicity in the cluster and the wide chemical dispersion, they conclude that Segue 1 is a disrupted dwarf spheroidal galaxy. Although the data presented here indicates the Orphan Stream progenitor is a disrupted dwarf spheroidal galaxy, we cannot reliably associate Segue 1 as the parent without additional observational data.

@@

The K-giants presented here can provide great insight
\begin{minipage}{\columnwidth}
into the chemistry and history of the Orphan Stream. High-resolution spectroscopic observations have been taken for some of our highly probable members and a detailed chemical analysis will be presented in a forthcoming paper (Casey et al., in preparation). Detailed chemical abundances can help determine both the nature of the progenitor before it is discovered, and allows us to compare peculiar chemical signatures with those of the known Milky Way satellites in order to associate likely parents. However, at least for the moment, the Orphan Stream remains appropriately named.
\end{minipage}

\acknowledgements
The authors would like to thank the anonymous referee for their constructive comments which improved this paper. ARC acknowledges the financial support through the Australian Research Council Laureate Fellowship 0992131, and from the Australian Prime Minister's Endeavour Award Research Fellowship, which has facilitated his research at MIT. SK and GDaC acknowledge the financial support from the Australian Research Council through Discovery Programs DP0878137 and DP120101237. GDaC is also grateful for the support received during an extended visit to the Institute of Astronomy, University of Cambridge, during which this work was completed.

Funding for the SDSS and SDSS-II has been provided by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S. Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England. The SDSS Web Site is http://www.sdss.org/.

The SDSS is managed by the Astrophysical Research Consortium for the Participating Institutions. The Participating Institutions are the American Museum of Natural History, Astrophysical Institute Potsdam, University of Basel, University of Cambridge, Case Western Reserve University, University of Chicago, Drexel University, Fermilab, the Institute for Advanced Study, the Japan Participation Group, Johns Hopkins University, the Joint Institute for Nuclear Astrophysics, the Kavli Institute for Particle Astrophysics and Cosmology, the Korean Scientist Group, the Chinese Academy of Sciences (LAMOST), Los Alamos National Laboratory, the Max-Planck-Institute for Astronomy (MPIA), the Max-Planck-Institute for Astrophysics (MPA), New Mexico State University, Ohio State University, University of Pittsburgh, University of Portsmouth, Princeton University, the United States Naval Observatory, and the University of Washington.

\vspace{2em}

\begin{thebibliography}{29}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Abazajian} {et~al.}(2009){Abazajian}, {Adelman-McCarthy},
  {Ag{\"u}eros}, {Allam}, {Allende Prieto}, {An}, {Anderson}, {Anderson},
  {Annis}, {Bahcall}, \& et~al.}]{Abazajian_et-al_2009}
{Abazajian}, K.~N., {et~al.} 2009, \apjs, 182, 543

\bibitem[{{Armandroff} \& {Da Costa}(1991)}]{Amandroff_Da_Costa_1991}
{Armandroff}, T.~E., \& {Da Costa}, G.~S. 1991, \aj, 101, 1329

\bibitem[{{Battaglia} \& {Starkenburg}(2012)}]{Battaglia_Starkenburg_2012}
{Battaglia}, G., \& {Starkenburg}, E. 2012, \aap, 539, A123

\bibitem[{{Belokurov} {et~al.}(2006){Belokurov}, {Zucker}, {Evans}, {Gilmore},
  {Vidrih}, {Bramich}, {Newberg}, {Wyse}, {Irwin}, {Fellhauer}, {Hewett},
  {Walton}, {Wilkinson}, {Cole}, {Yanny}, {Rockosi}, {Beers}, {Bell},
  {Brinkmann}, {Ivezi{\'c}}, \& {Lupton}}]{Belokurov_et-al_2006}
{Belokurov}, V., {et~al.} 2006, \apjl, 642, L137

\bibitem[{{Belokurov} {et~al.}(2007){Belokurov}, {Evans}, {Irwin},
  {Lynden-Bell}, {Yanny}, {Vidrih}, {Gilmore}, {Seabroke}, {Zucker},
  {Wilkinson}, {Hewett}, {Bramich}, {Fellhauer}, {Newberg}, {Wyse}, {Beers},
  {Bell}, {Barentine}, {Brinkmann}, {Cole}, {Pan}, \&
  {York}}]{Belokurov_et-al_2007}
---. 2007, \apj, 658, 337

\bibitem[{{Casey} {et~al.}(2012){Casey}, {Keller}, \& {Da
  Costa}}]{Casey_et-al_2012}
{Casey}, A.~R., {Keller}, S.~C., \& {Da Costa}, G. 2012, \aj, 143, 88

\bibitem[{{Fellhauer} {et~al.}(2007){Fellhauer}, {Evans}, {Belokurov},
  {Zucker}, {Yanny}, {Wilkinson}, {Gilmore}, {Irwin}, {Bramich}, {Vidrih},
  {Hewett}, \& {Beers}}]{Fellhaur_et-al_2007}
{Fellhauer}, M., {et~al.} 2007, \mnras, 375, 1171

\bibitem[{{Frebel} \& {Norris}(2011){Frebel}, \& {Norris}}]{Frebel_Norris_2011}
{Frebel}, A., \& Norris, J. E. 2011, arXiv/1102.1748

\bibitem[{{Grillmair} (2006)}]{Grillmair_2006}
{Grillmair}, C.~J. 2006, \apj, 643, L37

\bibitem[{{Hodapp} {et~al.}(2004){Hodapp}, {Kaiser}, {Aussel}, {Burgett},
  {Chambers}, {Chun}, {Dombeck}, {Douglas}, {Hafner}, {Heasley}, {Hoblitt},
  {Hude}, {Isani}, {Jedicke}, {Jewitt}, {Laux}, {Luppino}, {Lupton}, {Maberry},
  {Magnier}, {Mannery}, {Monet}, {Morgan}, {Onaka}, {Price}, {Ryan},
  {Siegmund}, {Szapudi}, {Tonry}, {Wainscoat}, \&
  {Waterson}}]{Hodapp_et-al_2004}
{Hodapp}, K.~W., {et~al.} 2004, Astronomische Nachrichten, 325, 636

\bibitem[{{Jin} \& {Lynden-Bell}(2007)}]{Jin_Lynden_Bell_2007}
{Jin}, S., \& {Lynden-Bell}, D. 2007, \mnras, 378, L64

\bibitem[{{Jordi} {et~al.}(2006){Jordi}, {Grebel}, \&
  {Ammon}}]{Jordi_et-al_2006}
{Jordi}, K., {Grebel}, E.~K., \& {Ammon}, K. 2006, \aap, 460, 339

\bibitem[{{Keller} {et~al.}(2007){Keller}, {Schmidt}, {Bessell}, {Conroy},
  {Francis}, {Granlund}, {Kowald}, {Oates}, {Martin-Jones}, {Preston},
  {Tisserand}, {Vaccarella}, \& {Waterson}}]{Keller_et-al_2007}
{Keller}, S.~C., {et~al.} 2007, PASA, 24, 1

\bibitem[{{Kerr} \& {Lynden-Bell}(1986)}]{Kerr_Lynden-Bell_1986}
{Kerr}, F.~J., \& {Lynden-Bell}, D. 1986, \mnras, 221, 1023

\bibitem[{{Koposov} {et~al.}(2010){Koposov}, {Rix}, \&
  {Hogg}}]{Koposov_et-al_2010}
{Koposov}, S.~E., {Rix}, H.-W., \& {Hogg}, D.~W. 2010, \apj, 712, 260

\bibitem[{{Marigo} {et~al.}(2008){Marigo}, {Girardi}, {Bressan}, {Groenewegen},
  {Silva}, \& {Granato}}]{Girardi_et-al_2008}
{Marigo}, P., {Girardi}, L., {Bressan}, A., {Groenewegen}, M.~A.~T., {Silva},
  L., \& {Granato}, G.~L. 2008, \aap, 482, 883

\bibitem[{{Mihalas} \& {Binney}(1981)}]{Mihalas_Binney_1981}
{Mihalas}, D., \& {Binney}, J. 1981, Science, 214, 829

\bibitem[{{Newberg} {et~al.}(2010){Newberg}, {Willett}, {Yanny}, \&
  {Xu}}]{Newberg_et-al_2010}
{Newberg}, H.~J., {Willett}, B.~A., {Yanny}, B., \& {Xu}, Y. 2010, \apj, 711,
  32

\bibitem[{{Odenkirchen} {et~al.}(2009){Odenkirchen}, {Grebel}, {Kayser}, {Rix},
  \& {Dehnen}}]{Odenkirchen_et-al_2009}
{Odenkirchen}, M., {Grebel}, E.~K., {Kayser}, A., {Rix}, H.-W., \& {Dehnen}, W.
  2009, \aj, 137, 3378

\bibitem[{{Robin} {et~al.}(2003){Robin}, {Reyl{\'e}}, {Derri{\`e}re}, \&
  {Picaud}}]{Robin_et-al_2003}
{Robin}, A.~C., {Reyl{\'e}}, C., {Derri{\`e}re}, S., \& {Picaud}, S. 2003,
  \aap, 409, 523

\bibitem[{{Roeser} {et~al.}(2010){Roeser}, {Demleitner}, \&
  {Schilbach}}]{Roeser_et-al_2010}
{Roeser}, S., {Demleitner}, M., \& {Schilbach}, E. 2010, \aj, 139, 2440

\bibitem[{{Sales} {et~al.}(2008){Sales}, {Helmi}, {Starkenburg}, {Morrison},
  {Engle}, {Harding}, {Mateo}, {Olszewski}, \& {Sivarani}}]{Sales_et-al_2008}
{Sales}, L.~V., {et~al.} 2008, \mnras, 389, 1391

\bibitem[{{Saviane} {et~al.}(2012){Saviane}, {Da Costa}, {Held}, {Sommariva},
  {Gullieuszik}, {Barbuy}, \& {Ortolani}}]{Saviane_et-al_2012}
{Saviane}, I., {Da Costa}, G.~S., {Held}, E.~V., {Sommariva}, V.,
  {Gullieuszik}, M., {Barbuy}, B., \& {Ortolani}, S. 2012, \aap, 540, A27

\bibitem[{{Schlegel} {et~al.}(1998){Schlegel}, {Finkbeiner}, \&
  {Davis}}]{Schlegel_Finkbeiner_Davis_1998}
{Schlegel}, D.~J., {Finkbeiner}, D.~P., \& {Davis}, M. 1998, \apj, 500, 525

\bibitem[{{Simon} {et~al.}(2011){Simon}, {Geha}, {Minor}, {Martinez},
	{Kirby}, {Bullock}, {Kaplinghat}, 
	{Strigari}, {Willman}, {Choi}, {Tollerud}, \&
	{Wolf}, J.}]{Simon_et-al_2011}
{Simon}, J.~D., {et~al.} 2011, \apj, 733, 46S

\bibitem[{{Starkenburg} {et~al.}(2010){Starkenburg}, {Hill}, {Tolstoy},
  {Gonz{\'a}lez Hern{\'a}ndez}, {Irwin}, {Helmi}, {Battaglia}, {Jablonka},
  {Tafelmeyer}, {Shetrone}, {Venn}, \& {de Boer}}]{Starkenburg_et-al_2010}
{Starkenburg}, E., {et~al.} 2010, \aap, 513, A34

\bibitem[{{Xue} {et~al.}(2008){Xue}, {Rix}, {Zhao}, {Re Fiorentin}, {Naab},
  {Steinmetz}, {van den Bosch}, {Beers}, {Lee}, {Bell}, {Rockosi}, {Yanny},
  {Newberg}, {Wilhelm}, {Kang}, {Smith}, \& {Schneider}}]{Xue_et-al_2008}
{Xue}, X.~X., {et~al.} 2008, \apj, 684, 1143

\bibitem[{{Zucker} {et~al.}(2006){Zucker}, {Belokurov}, {Evans}, {Kleyna},
  {Irwin}, {Wilkinson}, {Fellhauer}, {Bramich}, {Gilmore}, {Newberg}, {Yanny},
  {Smith}, {Hewett}, {Bell}, {Rix}, {Gnedin}, {Vidrih}, {Wyse}, {Willman},
  {Grebel}, {Schneider}, {Beers}, {Kniazev}, {Barentine}, {Brewington},
  {Brinkmann}, {Harvanek}, {Kleinman}, {Krzesinski}, {Long}, {Nitta}, \&
  {Snedden}}]{Zucker_et-al_2006}
{Zucker}, D.~B., {et~al.} 2006, \apjl, 650, L41

\end{thebibliography}

\end{document}

>>>>

\chapter{Hunting the Parent of the Orphan Stream}

% you can put a brief note at the top of your chapters using '\previouslypublished{}'. For example, to give a reference to the paper this chapter is based on, or to point out that it has been submitted to a journal.

\previouslypublished{Parts of this chapter have been previously published as \href{http://adsabs.harvard.edu/abs/2013ApJ...764...39C}{`Hunting the Parent of the Orphan Stream: Identifying Stream Members from Low-Resolution Spectroscopy', {Casey}, A.~R., {Da Costa}, G., {Keller}, S.~C., {Maunder}, E., 2013, ApJ, 764, 39C}. The work is presented here in expanded and updated form.}

\section{Introduction}
We present candidate K-giant members in the Orphan Stream which have been identified from low-resolution data taken with the AAOmega spectrograph on the Anglo-Australian Telescope. From modest S/N spectra and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. We find a revised stream distance of $22.5\,\pm\,2.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 82.1\,\pm\,1.4$\,km s$^{-1}$. The observed velocity dispersion of our most probable members is consistent with arising from the velocity uncertainties alone. This indicates that at least along this line-of-sight, the Orphan Stream is kinematically cold. Our data indicates an overall stream metallicity of [Fe/H] $= -1.63\,\pm\,0.19$\,dex which is more metal-rich than previously found and unbiased by spectral type. Furthermore, the significant metallicity dispersion displayed by our most probable members, $\sigma(\mbox{[Fe/H]}) = 0.56$\,dex, suggests that the unidentified Orphan Stream parent is a dSph satellite. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

\section{Introduction}
\label{sec:ch2-introduction}

@@

In this paper we present a detailed, self-consistent analysis to identify K-giant members of the Orphan Stream. Using our selection method we have catalogued the locations of nine highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:ch2-observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:ch2-analysis} and in \S\ref{sec:ch2-conclusions} the conclusions, predictions and future work are presented.

\section{Target Selection}
\label{sec:ch2-target-selection}

@@

\section{Observations}
\label{sec:ch2-observations}

@@

\section{Analysis \& Discussion}
\label{sec:ch2-analysis}

@@

Figure \ref{fig:ch2-velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin_et-al_2003}. We have selected particles from the Besan\c{c}on model using the same criteria outlined in \S\ref{sec:ch2-target-selection} after employing the \citet{Jordi_et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$.

@@

In a nearby ($\Delta\Lambda_{Orphan} \sim 4\,^\circ$) region of the stream, \citet{Newberg_et-al_2010} detected the Orphan Stream with a $V_{GSR} = 101.4$ km s$^{-1}$ from BHB stars. Differences in accounting for the local standard of rest between this work and \citet{Newberg_et-al_2010} means that this corresponds to approximately 95\,km s$^{-1}$ on our $V_{GSR}$ scale.  This is discussed further in \S\ref{sec:ch2-newberg}. The expected Orphan Stream kinematic peak is labelled in Figure \ref{fig:ch2-velocities}. There is no obvious sharp kinematic peak representative of the Orphan Stream in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution. To isolate potential Orphan Stream members we have nominated a relatively wide selection criteria between $65 \leq V_{GSR} \leq 125$ km s$^{-1}$ (shown in Figure \ref{fig:ch2-velocities}), which yields 28 Orphan Stream candidates. The typical uncertainty in our velocities is $\pm{}5.0$\,km s$^{-1}$.

\subsection{Dwarf/Giant Discrimination}
\label{sec:ch2-dwarf-giant}

@@

Figure \ref{fig:ch2-ew-mg} illustrates the trend with $EW_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel_Finkbeiner_Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. Giant stars populate the lower, sparser branch. A separation line has been adopted to distinguish dwarfs from giants, and is shown in Figure \ref{fig:ch2-ew-mg}. If we were to place this line higher, the total number of true giant stars may increase, but the dwarf contamination rate will rise dramatically. A compromise must be made between the rate of giant recoverability and the dwarf contamination. Our  dwarf/giant separation line lies just below the main dwarf population. On its own, this dwarf/giant separation line would typically result in far too many dwarf contaminants. However, we are  employing selections on multiple observables (kinematics, metallicity, proper motions) in order to refine our Orphan Stream giant sample.

@@

Our dwarf/giant separation line in Figure \ref{fig:ch2-ew-mg} yields 425 potential giants. Upon taking the intersection of our kinematic and gravity selections, we find 20 stars that appear to be likely Orphan Stream giants.

\subsection{Metallicities}
\label{sec:ch2-metallicities}

@@

Given a distance estimate to the Orphan Stream, we can also deduce a star's metallicity through isochrone fitting. We have used a 10\,Gyr \citet{Girardi_et-al_2008} isochrone at 21.4\,kpc \citep{Newberg_et-al_2010} and found metallicities for all 20 likely stream members from their best-fitting isochrone. Derived metallicities from Ca\,\textsc{II} line strengths and isochrone fitting that are consistent (within $\pm0.3$\,dex) indicates these measurements are reliable, and that these stars are indeed at a distance of $\sim$21.4\,kpc. We find ten highly likely stream members with consistently derived metallicities. They fall within the shaded region illustrated in Figure \ref{fig:ch2-feh}.

@@

Since our Orphan Stream giants cover a wide evolutionary range along the giant branch (Figure \ref{fig:ch2-cmd}), we are in a good position to revise the distance estimate to the stream. Given a 10\,Gyr \citet{Girardi_et-al_2008} isochrone at $[\mbox{Fe/H}] = -1.63$, we find a best-fitting distance to the stream of $22.5 \pm 2.0$\,kpc at $(l, b) = (250\,^\circ,\,50\,^\circ)$. This isochrone is shown in Figure \ref{fig:ch2-cmd}. Our derived distance is in reasonably good agreement with the measurement of $21.4 \pm 1.0$\,kpc independently deduced by \citet{Grillmair_2006} and \citet{Newberg_et-al_2010}.

\subsection{Comparison with \citet{Newberg_et-al_2010}}
\label{sec:ch2-newberg}
\citet{Newberg_et-al_2010} traced the Orphan Stream using BHB stars selected from the SEGUE survey, allowing them to derive an orbit for the stream and make a strong prediction for the location of the undiscovered progenitor. Their closest stream detection to this study is at $\Lambda_{Orphan} = 18.4\,^\circ$, approximately $\Delta\Lambda_{Orphan} \sim 4\,^\circ$ away from our fields. At this location, \citet{Newberg_et-al_2010} found the velocity of the stream to be $V_{GSR} = 101.4 \pm 8.9$\,km s$^{-1}$ based on 12 BHB stars. We note that this is $\sim95$\,km s$^{-1}$ on our scale, given the differences in accounting for the local standard of rest. The velocities and metallicities of our `High' and `Medium' probability candidates are illustrated in Figure \ref{fig:ch2-vgsr-feh}. Although we recover some candidates with velocities up to $V_{GSR} \sim 110$\,km s$^{-1}$, our kinematic distribution peaks near $V_{GSR} \sim 85$\,km s$^{-1}$, roughly 10\,km s$^{-1}$ lower than that of \citet{Newberg_et-al_2010}.

@@

The adopted metallicities of our Orphan Stream candidates are generally higher than those found by \citet{Newberg_et-al_2010}. Our highly likely stream members have a mean metallicity of [Fe/H] = $-1.63$, with a dispersion of $\sigma = 0.56$\,dex. As illustrated in Figures \ref{fig:ch2-vgsr-feh} and \ref{fig:ch2-newberg-feh}, there are two very metal-poor candidates which largely drive this dispersion, but we have no reason to suspect they are non-members. The \citet{Newberg_et-al_2010} sample contains 37 BHB stars identified over a 60\,$^\circ$ arc on the sky, and has a peak metallicity at [Fe/H] = $-2.10 \pm 0.10$. The closest detection bin in the \citet{Newberg_et-al_2010} sample was the most populous, yielding 7 BHB stars. For comparison, we identify 9 giant stars across $\sim{}4\,^\circ$. Given BHB stars are known to trace a somewhat more metal-poor population, and we are calculating statistics with marginal sample sizes, we conclude that the accuracy of these two metallicity distributions are not mutually exclusive. It is entirely possible that we are sampling the same distribution, but a larger sample size is required.

\section{Conclusions}
\label{sec:ch2-conclusions}

@@

Given the stream orbit derived by \citet{Newberg_et-al_2010}, they excluded all possible known halo objects except for the dissolved star cluster, Segue 1. \citet{Simon_et-al_2011} obtained spectroscopy for six members in Segue 1 and found an extremely wide metallicity dispersion: from $<-3.4$ to $-1.63$ dex. On the basis of the extremely low metallicity in the cluster and the wide chemical dispersion, they conclude that Segue 1 is a dwarf spheroidal galaxy. Although the data presented here indicates the Orphan Stream progenitor is a disrupted dwarf spheroidal galaxy, we cannot reliably associate Segue 1 as the parent without additional observational data.

@@

The K-giants presented here can provide great insight into the chemistry and history of the Orphan Stream. High-resolution spectroscopic observations have been taken for some of our highly probable members and a detailed chemical analysis will be presented in a forthcoming paper (Casey et al., in preparation). Detailed chemical abundances can help determine both the nature of the progenitor before it is discovered, and allows us to compare peculiar chemical signatures with those of the known Milky Way satellites in order to associate likely parents. However, at least for the moment, the Orphan Stream remains appropriately named.
